### 2015-04-16 ind-steelsbook.tex

\documentclass[ number=1
			   ,series=cmle
			   ,url=http://langsci-press.org/ 
			   ,isbn=978-3-944675-42-8
			   ,output=short   % long|short|inprep              
			   %,blackandwhite
			   ,smallfont
			   ,biblatex
			   ,draftmode  
			   %,draft
			  ]{LSP/langsci}                          
\usepackage{LSP/lsp-styles/lsp-gb4e}		% verhindert Komma bei mehrfachen Fußnoten?
\usepackage{LSP/lsp-styles/avm}  
\usepackage{todonotes}                                                  
\usepackage{layout}  
\usepackage{lipsum}   
% \usepackage{covington}
% \usepackage{tipa}
% \usepackage[T1]{fontenc}
% \usepackage{tipx}
\usepackage{qtree}
\usepackage{amsmath}
\usepackage{multirow} 
\usepackage{multicol}
%\usepackage[latin1]{inputenc}
\usepackage{authorindex}
%\usepackage[round]{natbib}

\usepackage{localcommands}
\usepackage{localhyphenation}

\title{The Talking Heads Experiment}
\subtitle{Origins of Words and Meanings}
\BackTitle{The Talking Heads Experiment}% - \\ Origins of Words and Meanings}
\author{Luc Steels}
%\Affiliation{ Catalan Institute for Advanced Studies (ICREA) \\ %- Instituci\accute{o} Catalana de Recerca i Estudis Avan\cats\\
%Institut de Biologia Evolutiva (Universitat Pompeu Fabra and CSIC) Barcelona}
\dedication{For my daughter Lenie}
\BackBody{The Talking Heads Experiment, conducted in the years 1999--2001, was the first large-scale experiment 
in which open populations of situated embodied agents created for the first time ever a new shared vocabulary
by playing language games about real world scenes in front of them. The agents could teleport to different 
physical sites in the world through the Internet. Sites, in Antwerp, Brussels, Paris, Tokyo, London, Cambridge
and several other locations were linked into the network. Humans could interact with the robotic agents either
on site or remotely through the Internet and thus influence the evolving ontologies and languages of the 
artificial agents.

The experiment tested theories of the origins of language and meaning and of the grounding of language in 
real world perception and action. The agents have been endowed with a cognitive architecture that enables them to 
invent their own ontologies and languages and acquire those already existing in the group. Biological principles
such as self-organisation and selectionism played a major role to attain the emergence and continued evolution of 
this agent-based information ecosystem.}

\bibliography{references} % References file
\begin{document}

\maketitle                
\frontmatter
\include{preface/preface}
\tableofcontents

\mainmatter

%\part{The 1999 Talking Heads Book} \label{part:1}
%
%%\thispagestyle{empty}
%\begin{figure}[htbp]
  %\centerline{\includegraphics[width=1.0\textwidth]{extra/figs/cover-original-book.pdf}}
%\caption{\small\label{left-right} Original cover of the Talking Heads book published in 1999 on the occasion of the 
%Laboratorium exhibition in Antwerp.}
%\label{f:cover}
%\end{figure}

%\include{chap1/chap1}
%\include{chap2/chap2}
%\include{chap3/chap3}
%\include{chap4/chap4}
%\include{chap5/chap5}
%\include{chap6/chap6}
%\include{chap7/chap7}
%%
%\part{Installations and experiments} \label{part:2}
%\include{chap8/chap8}
%\include{chap9/chap9}
%
\part{Beyond the Talking Heads} \label{part:3}
\include{chap10/chap10}
\include{chap11/chap11}
\include{chap12/chap12}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%             Backmatter                       %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter

\part{Appendices} \label{part:4}
\include{extra/papers-language-steels}

\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{chapter}{List of references}
\printbibliography %change to the name of your bib file
\clearpage

\addcontentsline{toc}{chapter}{Index} 
\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{section}{Name index}
\ohead{Name index} 
\printindex

\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{section}{Language index}
\ohead{Language index} 
\printindex[lan]

\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{section}{Subject index}
\ohead{Subject index} 
\printindex[sbj]

\end{document}